# GDP Uncertainty Between 1997 and 2007

## Visualizing Uncertainty


This analysis uses uncertainty-based graphics to reveal global income disparities and their impact on life expectancy. Faceted histograms show GDP distributions, while density plots link economic growth to rising life expectancy. Despite persistent gaps, higher median wealth correlates with longer lifespans. Plotting uncertainty enhances pattern comprehension, aiding policymakers in addressing health and economic inequalities.


```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)  # For ggplot, dplyr, and friends
library(gapminder)  # For health/wealth data
library(ggridges)   # For ridge plots
library(gghalves)   # For half geoms
library(hrbrthemes)

```

## Histograms

```{r filter-gapminder-1997}
# Create a filtered dataset
gapminder_1997 <- gapminder %>%
  filter(year == 1997)

# View the filtered dataset
print(gapminder_1997)
```



```{r plot-histogram}
# plot-histogram
ggplot(gapminder_1997, aes(x = log(gdpPercap), fill = continent)) +
  geom_histogram(binwidth = 0.2, alpha = 0.8, color = NA) +  # Remove black outlines
  labs(
    title = "Histogram of Logged GDP per Capita (1997)",
    x = "Log(GDP per Capita)",
    y = "Frequency",
    fill = "Continent"
  ) +
  scale_fill_viridis_d() +  # Use viridis discrete color scale
  theme_ipsum(base_family = "sans") +  # Set font to default system font
  theme(
    axis.text.x = element_text(size = 10, color = "#585858", 
                               margin = margin(t = 10)),
    axis.text.y = element_text(size = 12, color = "#000", 
                               margin = margin(r = 10)),
    axis.title.x = element_text(size = 12, color = "#202d53", 
                                margin = margin(t = 10), hjust = 0.5),
    axis.title.y = element_text(size = 12, color = "#202d53", 
                                margin = margin(r = 10), hjust = 0.5),
    plot.margin = margin(10, 15, 15, 10),
    legend.position = "right",
    plot.title = element_text(color = "darkblue", size = 16, hjust = 0.5)
  )
```


```{r histogram with facets, warning=FALSE, message=FALSE}
# Define custom colors using viridis palette
library(viridis)

# Create a faceted histogram with log-transformed GDP per capita
ggplot(gapminder_1997, aes(x = gdpPercap, fill = continent)) +
  geom_histogram(binwidth = 0.17, alpha = 0.8, color = NA) +  # Remove black outlines
  scale_x_log10(labels = function(x) scales::trans_format("log10", 
                                                          scales::number_format(scale = 1))(x)) +  
  # Scale and simplify labels
  facet_wrap(~continent) +  # Create facets for each continent
  labs(
    title = "Histogram of Logged GDP per Capita (1997) by Continent",
    x = "GDP per Capita (Log Scale)",
    y = "Frequency",
    fill = "Continent"
  ) +
  scale_fill_viridis_d() +  # Use the viridis color palette
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, color = "#585858",
                               margin = margin(t = 10)),  # Adjust x-axis label color
    axis.text.y = element_text(size = 10, color = "#000",
                               margin = margin(r = 10)),  # Adjust y-axis label color
    axis.title.x = element_text(size = 12, color = "#202d53",
                                margin = margin(t = 10)),  # Adjust x-axis title color
    axis.title.y = element_text(size = 12, color = "#202d53",
                                margin = margin(r = 10)),  # Adjust y-axis title color
    plot.margin = margin(10, 15, 15, 10),  # Adjust margins
    legend.position = "none",  # Remove legend since facets have labels
    plot.title = element_text(color = "darkblue", size = 16, hjust = 0.5),  
    # Change title color to dark blue
    panel.spacing = unit(1.5, "lines")  # Increase space between facets
  )

```


The filtered 1997 GDP per Capita dataset shows left skewness with Africa dominantly. In contrast, the income were more abundant and concentrated to the right for Europe and Asia, except that Asia had a wider range of distribution. America had almost the same range of GDP but they were more skewed to the right showing that only a certain number of people around less than 2 million could make more than five figures. On the other hand, there was a noticeable limited variance for Oceania in which a majority of them were among the highest scale of GDP per Capita. The first histogram is scaled by log from 6, standing for $1 milliion (if the units are in USD).



## Densities

```{r plot-ridgeplot, warning=FALSE, message=FALSE}

ggplot(gapminder, aes(x = lifeExp, y = factor(year), fill = factor(year))) +
  geom_density_ridges(quantile_lines = TRUE, quantiles = 2, 
                      binwidth = 0.2, alpha = 0.8, scale = 1.2) +  
  # Increase scale for more space between distributions
  scale_x_continuous(breaks = seq(20, 90, by = 10)) +  
  # Set breaks at intervals of 10 for clarity
  scale_fill_viridis_d() +  # Use a color palette
  labs(
    title = "Ridge Plot of Life Expectancy Over Time (1952-2007)",
    x = "Life Expectancy",
    y = "Year",
    fill = "Year"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, color = "#585858", margin = margin(t = 2)),
    axis.text.y = element_text(size = 10, color = "#000", margin = margin(r = 1)),
    axis.title.x = element_text(size = 12, color = "#202d53", margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, color = "#202d53", margin = margin(r = 10)),
    plot.margin = margin(10, 15, 15, 10),
    plot.title = element_text(color = "darkblue", size = 16, hjust = 0.5)
  )


```


Aside from golbal wealth inspection, one can lay out some association to health. For example, Density plot is helpful in handling varied sets of distribution in quantiles, such as the life expectancy of each year. From 1952 to 2007, there was an upward trend of life expectancy with the median shifting towards the 75th quantile each year, indicating less than 50% of the population lived up to over 65 years since 1987. The trend also became more uniformed as time progressed instead of bimodal; hence, less variability in reaching older ages. One can assume that though Africa was the greatest weight of still developing individual gross production in 1997, the combined global wealth of higher scale of other contients seemed to lead to better access to healthcare.
 
## GDP per Capita in The Next Ten Years9


```{r filter-gapminder-2007-sans-oceania}
# Create a filtered dataset

gapminder_2007 <- gapminder %>%
  filter(year == 2007 & continent != "Oceania")

# Show the filtered dataset
print(gapminder_2007)

```



```{r plot-distribution, warning=FALSE, message=FALSE}
# Create a combined half-boxplot and half-violin plot with log scale
ggplot(gapminder_2007, aes(x = continent, y = gdpPercap, fill = continent)) +
  geom_half_violin(side = "l", alpha = 0.6, color = NA) +  # Left-side violin plots
  geom_half_boxplot(side = "r", width = 0.2, outlier.shape = NA, color = "black", alpha = 0.8) +  
  # Right-side boxplots
  geom_jitter(width = 0.1, size = 1.5, alpha = 0.6) +  
  # Add jittered points for raw data visualization
  scale_y_log10(labels = scales::label_comma()) +  
  # Apply log transformation on y-axis
  labs(
    title = "Distribution of GDP per Capita (2007, Log Scale)",
    x = "Continent",
    y = "GDP per Capita (Log Scale)"
  ) +
  scale_fill_viridis_d() +  # Use Viridis for better contrast
  theme_minimal(base_size = 16) +  # Improve readability
  theme(
    text = element_text(family = "sans"),  # Set a safe default font to avoid errors
    legend.position = "none",  # Remove legend
    plot.title = element_text(size = 15, hjust = 0.5, color = "darkblue"),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10, color = "#585858"),
    axis.title.x = element_text(size = 12, color = "#202d53", margin = margin(t = 8)),
    axis.title.y = element_text(size = 12, color = "#202d53"),
    panel.grid.major = element_line(color = "#d3d3d3", size = 0.5)  # Light grid lines
  )
```

It was seen that in 2007, more than 50% of the population lived over 70 years of age; assuming people were financially making more to support and excel the basic needs for life. In 1997, a great majority of Africans produced four figures and less per Capita between $350 to $2,000 especially, and that concentration was spread out abit with the lowest global median of $1,700 in 2007 compared to 1997 around $1,300. The median of wealth for each continent was increased between 10 to 30 percent while the spread were becoming more clustered. Regardless the change, Europe was still considered rich in GDP per Capita than other continents with the median reaching $30,000. Thus, small incremental changes to the gap of wealth caused the world to achieve greater life expectancy.

```{r Filter-for-1997}
# Create a filtered dataset for 1997

gapminder_1997 <- gapminder %>%
  filter(year == 1997 & continent != "Oceania")

# Show the filtered dataset
print(gapminder_1997)
```

```{r plot-distribution 1997, warning=FALSE, message=FALSE }
# Create a combined half-boxplot and half-violin plot with log scale for 1997
ggplot(gapminder_1997, aes(x = continent, y = gdpPercap, fill = continent)) +
  geom_half_violin(side = "l", alpha = 0.6, color = NA) +  # Left-side violin plots
  geom_half_boxplot(side = "r", width = 0.2, outlier.shape = NA, color = "black", alpha = 0.8) +  
  # Right-side boxplots
  geom_jitter(width = 0.1, size = 1.5, alpha = 0.6) +  
  # Add jittered points for raw data visualization
  scale_y_log10(labels = scales::label_comma()) +  
  # Apply log transformation on y-axis
  labs(
    title = "Distribution of GDP per Capita (1997, Log Scale)",
    x = "Continent",
    y = "GDP per Capita (Log Scale)"
  ) +
  scale_fill_viridis_d() +  # Use Viridis for better contrast
  theme_minimal(base_size = 16) +  # Improve readability
  theme(
    text = element_text(family = "sans"),  # Set a safe default font to avoid errors
    legend.position = "none",  # Remove legend
    plot.title = element_text(size = 15, hjust = 0.5, color = "darkblue"),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10, color = "#585858"),
    axis.title.x = element_text(size = 12, color = "#202d53", margin = margin(t = 8)),
    axis.title.y = element_text(size = 12, color = "#202d53"),
    panel.grid.major = element_line(color = "#d3d3d3", size = 0.5)  # Light grid lines
  )
```
Disparities between continents are revealed by analyzing GDP per capita distributions, highlighting the necessity of accurately depicting variability. Density plots and faceted histograms demonstrate how financial stability enhances health outcomes by illuminating wealth concentration and its effect on life expectancy. The change in life expectancy distributions emphasizes less fluctuation over time, hence strengthening the connection between longevity and economic growth. Better policy decisions can be made by using uncertainty plotting to find patterns, outliers, and structural disparities.

